### 2011-02-22 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
final version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1279 d8b0c59b-6544-0410-94d9-83a5658a058e

This deliverable is structured as follows: 
we will start off by outlining the main changes that this document has undergone since its previous version, so as to enable the reader to quickly identify new material.
After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2010fast_state_of_the_art} --- the structure of the following sections follows directly from our adopted ontology development methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies. This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the concrete ontology in its form at this moment of the project, as well as some basic considerations such as an early proposal for prototypical semantics for RDF. 
After a section discussing an evaluation of the FAST gadget ontology in Sect.~\ref{sec:evaluation}, the deliverable will be concluded with a summary of our work in Sect.~\ref{sec:conclusions}.\ref{sub:methontology_a_methodology_for_ontology_development}
A number of appendices have been added to the document: App.~\ref{sec:methodology} presents the development methodology adopted by us, App.~\ref{sec:ontology_terms} provides documentation for the ontology in the form of a detailed list of terms, while App.~\ref{sec:ontology_code} contains the complete code of the ontology.

@@

\begin{itemize}
	\item Section~\ref{sub:classes_instances_and_templates} has been significantly extended with a more in-depth discussion of prototype theory, and a first proposal for prototype-based semantics for RDF, as they are employed in FAST.
	\item A section for the \emph{VocDoc} tool, which was developed in part for generating online and PDF documentation for the FAST ontology, has been added as App.~\ref{sec:vocdoc}.
	\item The ontology itself has remained more or less stable since D2.2.2.~\cite{moeller2010fast_ontology}. However, a handful of small changes and corrections have been performed, e.g., to reflect the terminology changes in the prototype section (\emph{prototype} instead of \emph{template}, \emph{clone} instead of \emph{copy}).
	\item Based on comments from the previous review, Fig.~\ref{fig:ontology_uml} has been corrected and improved with a legend.
	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been throughout the deliverable.
	\item The online documentation of the FAST Gadget ontology at \url{http://purl.oclc.org/fast/ontology/gadget} has been updated to reflect the changes in this version of the deliverable.
\end{itemize}

@@

\textbf{Properties:} \emph{containsScreen}, \emph{hasLabel}, \emph{hasIcon}, \emph{hasPreCondition}, \emph{hasPostCondition}, \emph{hasTag}, \emph{hasProfile}\\ \hline
\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{urena2010fast_architecture}, FAST Requirements Specification~\cite{villoslada2010fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
\end{longtable}
\end{small}
\doublespacing

@@

\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{urena2010fast_architecture}. In terms of RDF, a fact is a statement consisting of a subject, predicate and object \emph{(S, P, O)}. \\ \hline

@@

Regarding the layering from the actual gadget down to the embedded Web services, the central classes in the conceptual model of FAST can be grouped into three levels: the gadget and screen flow level at the top, the level of individual screens in the middle, and the level of Web services at the bottom. This layering is reflected in Fig.~\ref{fig:uml_composition}, which illustrates the central compositional relationships of the model. As the figure shows, a \emph{gadget} in FAST is mainly composed of a \emph{screen flow}, which it wraps. The \emph{screen flow} is composed of one or more \emph{screens}, which in turn comprises the three \emph{screen components} of the visual front-end (the \emph{form}), the back end service (the \emph{resource}) and the data \emph{operator}. Finally, \emph{forms} are assembled from \emph{form elements}, while \emph{resources} wrap \emph{backend services}.

@@

Figure~\ref{fig:uml_screen_component} looks closer at the features of individual screens. These are forms, resources and operators, which are all specialisations of the general \emph{screen component} concept. Each such screen component may or may not have an action or trigger associated with them, which declaratively define their behaviour in terms of their own functionality or functionality they can trigger in other building blocks.

@@

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/declarativeDefinitionExample.pdf}
    \caption{Explicit piping within a screen from a post- to a pre-condition (diagramme)}
    \label{fig:piping}
  \end{center}
\end{figure}

@@

>>>>

This deliverable is structured as follows: 
we will start off by outlining the main changes that this document has undergone since its previous version, so as to enable the reader to quickly identify new material.
After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2010fast_state_of_the_art} --- the structure of the following sections follows directly from our adopted ontology development methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies. This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the concrete ontology in its form at this moment of the project, as well as some basic considerations such as an early proposal for prototypical semantics for RDF. 
After a section discussing an evaluation of the FAST gadget ontology in Sect.~\ref{sec:evaluation}, the deliverable will be concluded with a summary of our work in Sect.~\ref{sec:conclusions}.\ref{sub:methontology_a_methodology_for_ontology_development}
A number of appendices have been added to the document: App.~\ref{sec:methodology} presents the development methodology adopted by us, App.~\ref{sec:ontology_terms} provides documentation for the ontology in the form of a detailed list of terms, while App.~\ref{sec:ontology_code} contains the complete code of the ontology. Appendix~\ref{sec:vocdoc} describes and documents the VocDoc tool, which was developed to automatically create ontology documentation from the original ontology code.

@@

\begin{itemize}
	\item Section~\ref{sub:classes_instances_and_templates} has been significantly extended with a more in-depth discussion of prototype theory, and a first proposal for prototype-based semantics for RDF, as they are employed in FAST.
	\item A section for the \emph{VocDoc} tool, which was developed in part for generating online and PDF documentation for the FAST ontology, has been added as App.~\ref{sec:vocdoc}.
	\item The ontology itself has remained more or less stable since D2.2.2.~\cite{moeller2010fast_ontology}. However, a handful of small changes and corrections have been performed, e.g., to reflect the terminology changes in the prototype section (\emph{prototype} instead of \emph{template}, \emph{clone} instead of \emph{copy}).
	\item Based on comments from the previous review, Fig.~\ref{fig:ontology_uml} has been corrected and improved with a legend.
	\item Also in response to reviewer comments, some concrete examples for instances of the various classes in the FAST ontology have been added to Sect.~\ref{ssub:term_relations_within_the_conceptual_model}.
	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been made throughout the deliverable.
	\item The online documentation of the FAST Gadget ontology at \url{http://purl.oclc.org/fast/ontology/gadget} has been updated to reflect the changes in this version of the ontology.
\end{itemize}

@@

\textbf{Properties:} \emph{containsScreen}, \emph{hasLabel}, \emph{hasIcon}, \emph{hasPreCondition}, \emph{hasPostCondition}, \emph{hasTag}, \emph{hasProfile}\\ \hline
\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{lopez2011fast_architecture}, FAST Requirements Specification~\cite{villoslada2010fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
\end{longtable}
\end{small}
\doublespacing

@@

\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{lopez2011fast_architecture}. In terms of RDF, a fact is a statement consisting of a subject, predicate and object \emph{(S, P, O)}. \\ \hline

@@

Regarding the layering from the actual gadget down to the embedded Web services, the central classes in the conceptual model of FAST can be grouped into three levels: the gadget and screen flow level at the top, the level of individual screens in the middle, and the level of Web services at the bottom. This layering is reflected in Fig.~\ref{fig:uml_composition}, which illustrates the central compositional relationships of the model. As the figure shows, a \emph{gadget} in FAST is mainly composed of a \emph{screen flow}, which it wraps. The \emph{screen flow} is composed of one or more \emph{screens}, which in turn comprises the three \emph{screen components} of the visual front-end (the \emph{form}), the back end service (the \emph{resource}) and the data \emph{operator}. Finally, \emph{forms} are assembled from \emph{form elements}, while \emph{resources} wrap \emph{backend services}. To illustrate, Fig.~\ref{fig:screen} shows the picture of a screen, as it is actually rendered in a web browser.

@@

\begin{figure}
  \begin{center}
    \includegraphics[scale=.5]{images/screen.png}
    \caption{A FAST screen in a web browser}
    \label{fig:screen}
  \end{center}
\end{figure}

Figure~\ref{fig:uml_screen_component} looks closer at the features of individual screens. These are forms, resources and operators, which are all specialisations of the general \emph{screen component} concept. Each such screen component may or may not have an action or trigger associated with them, which declaratively define their behaviour in terms of their own functionality or functionality they can trigger in other building blocks. Some concrete examples of various different screen components are given in Fig.~\ref{fig:building_blocks}, which shows a screenshot from the screen composition tool in the FAST GVS \cite{lopez2011fast_architecture}.

@@

\begin{figure}
  \begin{center}
    \includegraphics[scale=.66]{images/building_blocks.pdf}
    \caption{Examples of FAST building blocks during screen composition}
    \label{fig:building_blocks}
  \end{center}
\end{figure}

@@

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\linewidth]{images/declarativeDefinitionExample.pdf}
    \caption{Explicit piping within a screen from a post- to a pre-condition (diagram)}
    \label{fig:piping}
  \end{center}
\end{figure}

### 2011-02-22 trunk/deliverables/D2_2_Ontology/properties.tex
final version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1279 d8b0c59b-6544-0410-94d9-83a5658a058e

@@

\subsubsection*{Property: \texttt{fgo:hasCopy}}
\label{subs:hasCopy}
\textcolor{white}{\textbf{label}} & has copy \\ \hline
\textcolor{white}{\textbf{description}} & Links the template of a building block as used in the palette or catalogue of available building blocks to its copy, as used in a particular screen flow. \\ \hline

@@

@@

\subsubsection*{Property: \texttt{fgo:hasTemplate}}
\label{subs:hasTemplate}
\textcolor{white}{\textbf{label}} & has template \\ \hline
\textcolor{white}{\textbf{description}} & Links the copy of a building block as used in a particular screen flow to its template, as used in the palette or catalogue of available building blocks. \\ \hline

>>>>

\subsubsection*{Property: \texttt{fgo:hasClone}}
\label{subs:hasClone}
\textcolor{white}{\textbf{label}} & has clone \\ \hline
\textcolor{white}{\textbf{description}} & Links the prototype of a building block as used in the palette or catalogue of available building blocks to its clone, as used in a particular screen flow or screen. \\ \hline
\textcolor{white}{\textbf{inverse\_of}} & \htmlref{\texttt{fgo:hasPrototype}}{subs:hasPrototype} \\ \hline

@@

@@

\subsubsection*{Property: \texttt{fgo:hasPrototype}}
\label{subs:hasPrototype}
\textcolor{white}{\textbf{label}} & has prototype \\ \hline
\textcolor{white}{\textbf{description}} & Links the clone of a building block as used in a particular screen flow or screen to its prototype, as used in the palette or catalogue of available building blocks. \\ \hline
\textcolor{white}{\textbf{inverse\_of}} & \htmlref{\texttt{fgo:hasClone}}{subs:hasClone} \\ \hline

@@
